\input{/Users/ubd/Library/Mobile Documents/iCloud~md~obsidian/Documents/rhizome/06_projects/UNI/latex_template}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title & Author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Title}
\author{Utku B. Demir}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\section{Mirroring representative claims and its surrogative expansion in
  non-electoral movements: Zapatistas and Rojava Resistance}

What is your essay question?

How do non-electoral movements, especially armed ones like the Zapatistas and the Rojava resistance, generate their representative claims aiming for a surrogate representation of some communities on global scale, and what lessons do they offer for democratic theory and practice?

Why do you think this question is important or interesting?

Activist groups generate representative claims in different forms. The
armed militant groups often are formed around a representative claim, in
case of Zapatistas for example Saward \parencite[13]{saward2010}
associates their struggle to make the reflect the voices of indigenous
people of Chiapas as \textit{mirroring claim}. However,  both Saward and
Tormey \parencite[]{tormey2006} do not reflect on the claims
Zapatistas are making for the global issues, struggles, other
communities. In the case of Zapatistas, they claim to be the voice of all
the oppressed, marginalised communities around the world (see Marcos'
letter and Zapatistas support on other struggles around the world), in
the case of Rojava the emphasis on women's, LGBT rights, and
intersectionalitly was especially especially prominent within the contrast against
the surrounding forces. On top of the symbolic claims, Rojava's claims
also found some direct references in other international movements (e.g.
jin jiyan azadi slogan in Persian movement against islamic regime). While
these forces claim to represent a specific communities' values, they are
also simultaneously claim representing global causes as well, I aim to
analyse the nature and effects of such claims on both fronts.




How do you define the representation conceptions you are engaging with?
Saward's formulation of the representative claims to identify the claims
among both movements, and Mansbridge concept of surrogative
representation to discuss, if and in what this concept can also contain
non-electoral representative claims.

What argument do you want to make?
Their mirroring representative claims is forming a constituency among
their local communities through the global values, and struggles the
organisations aim to be in solidarity to. This contructivist approach,
however, allow other global communities to form movements around their claims.

What examples, evidence or data do you want to use to support your argument ?
hI'll mostly refer to the hesources about the acthions of those movements

Do you have any specific things you are struggling with or where you would like advice by your peers?

\printbibliography
\end{document}
