\input{/Users/ubd/Library/Mobile Documents/iCloud~md~obsidian/Documents/rhizome/06_projects/UNI/latex_template}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title & Author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{20250320\_Meeting with Gorazd}
\author{Utku B. Demir}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\section{Matter with the CALL:HORIZON-CL2-2025-01-HERITAGE-04: Leveraging artificial intelligence for creativitydriven innovation. }

\begin{itemize}
	\item The call is canceled from the beginning.
	\item I have been excluded deliberately.
	\item The communication was led \textbf{indirectly}
	\item If some previous disputes like AI-Group is related, remains to be seen.
\end{itemize}

\section{BIO4Campus \& FOR-FREIGHT}

\begin{itemize}
	\item The question of Dienstreise.
	\item The matter of a really tight timeframe regarding BIO4Campus.
	\item Tasks regarding FF have to be dropped (Dietmar and others are to be
	      informed), including Project Meeting organisation.
\end{itemize}


\section{Calls I am interested in (CL2)}
\begin{itemize}
	\item HORIZON-CL2-2025-01-DEMOCRACY-06: Countering and preventing radicalisation, extremism, hate speech and polarisation.
	\item HORIZON-CL2-2025-01-DEMOCRACY-08: Understanding and addressing gender-based violence against politically active women and LGBTIQ people
	\item \hlred{HORIZON-CL2-2025-01-DEMOCRACY-10: Economic inequalities and their impact on democracy}
	\item \hlred{HORIZON-CL2-2025-01-DEMOCRACY-12: The role of civic and citizenship education for strengthening civic and democratic participation and support for common European values.}
	\item \hlred{HORIZON-CL2-2025-01-DEMOCRACY-15: Community of democracy practitioners and researchers}
	\item \hlred{HORIZON-CL2-2025-01-HERITAGE-04: Leveraging artificial intelligence for creativitydriven innovation}
	\item HORIZON-CL2-2025-01-TRANSFO-01: Tackling gender-based violence in different social and economic spheres
	\item HORIZON-CL2-2025-01-TRANSFO-05: Improving fairness in the economy through a better understanding of undeclared and underdeclared work
	\item HORIZON-CL2-2025-01-TRANSFO-07: Impact of the learning environment and the use  of digital tools in everyday life on key skills and competence development.
	\item \hlred{HORIZON-CL2-2025-01-HERITAGE-05: Evolution of culture in a virtualising world ... 89  INNOVATIVE}
\end{itemize}

\printbibliography
\end{document}
