\input{/Users/ubd/Library/Mobile Documents/iCloud~md~obsidian/Documents/rhizome/06_projects/UNI/latex_template}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Title & Author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Research Diary}
\author{Utku B. Demir}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle
\section*{Research Diary}

\subsection*{October: Laying the Groundwork}
The research project commenced in October with an introduction to actor and issue mapping as foundational tools for analyzing the debate surrounding aspartame. The initial focus was on constructing the \textbf{Actor Map}, which identified key stakeholders in the debate. These included global regulatory bodies such as the WHO and EFSA, industry leaders like Coca-Cola, advocacy groups, media outlets, and consumers. This exercise revealed intricate relationships, such as the influence of industry lobbying on regulatory decisions and the role of media in shaping public perception.

The next step involved creating the \textbf{Issue Map}, where we examined central controversies such as the potential carcinogenicity of aspartame, regulatory transparency, and public misinformation. This mapping process helped clarify how different actors engage with these issues and their broader societal implications. By the end of the month, both maps were completed, providing a clear analytical framework for the subsequent phases of the project.

\subsection*{November: Preparing for Engagement}
In November, the focus shifted to deepening our analysis and preparing for stakeholder interviews. The finalized \textbf{Interview Guides} were designed to target specific aspects of the aspartame debate, with tailored questions addressing scientific, societal, and regulatory dimensions. These guides reflected key insights from the actor and issue maps, ensuring the interviews would provide meaningful data.

While outreach efforts to societal actors faced delays, we successfully secured an interview with \textbf{Veronika Plichta} from AGES to represent the scientific perspective. Alongside this, we conducted extensive literature reviews, examining regulatory reports, scientific papers, and public advocacy materials. This process highlighted gaps in the discourse, such as the disconnect between public perception and scientific consensus.

\subsection*{December: Gathering Insights}
December was dedicated to conducting interviews and synthesizing the collected data. The interview with \textbf{Veronika Plichta} provided valuable insights into how regulatory bodies evaluate artificial sweeteners like aspartame. She emphasized the reliance on evidence-based thresholds and international safety standards while addressing public misconceptions fueled by media coverage.

Parallel to this, we compiled the findings from our actor and issue maps, interviews, and literature reviews into a comprehensive \textbf{Dossier}. This document served as a critical resource for the upcoming \textbf{Panel Discussion}, where we presented our findings and engaged with potential solutions to the controversies surrounding aspartame. This phase underscored the need for effective communication between scientific, regulatory, and societal actors.

\subsection*{January: Completing the Puzzle}
The final phase of the project in January focused on completing deliverables and integrating delayed inputs. After persistent outreach efforts, we secured an interview with \textbf{Ajda Ahmedova} from Fortrea, representing the societal perspective. Her contributions highlighted significant concerns about industry influence on regulatory processes and the erosion of public trust in scientific assessments. This interview added a much-needed societal dimension to our research.

The project culminated in the drafting of the \textbf{Policy Brief}, where we proposed actionable recommendations such as harmonizing global regulatory standards, enhancing transparency, and improving public communication. This document synthesized months of research and provided a roadmap for addressing the challenges associated with artificial sweeteners like aspartame. Reflecting on the journey, this phase demonstrated the importance of interdisciplinary collaboration and the balance between evidence and societal concerns.

\printbibliography
\end{document}
